---
title: "1_pds_threetimpts_cleaning_script"
author: "Jackie"
date: "2/15/2020"
output: html_notebook 
---

# loading libraries
```{r}
library(lmerTest)
library(tidyverse)
library(foreign)
library(haven)
library(readxl)
library(naniar)
```

# read in sex info
### sex coding  
_1 = Male_  
_2 = Female_
```{r}
demo_t1_fp <- "~/Box/demographics_golden_standard.T1.sav"
demo_t1 <- read_sav(demo_t1_fp) %>%
  mutate(
    ELS_ID = as.factor(ELS_ID)
    ) %>%
  dplyr::select(ELS_ID,
                Child_Sex.T1)
  
demo_t1_clean <- 
  demo_t1 %>%
  mutate(Child_Sex.T1 = as.factor(Child_Sex.T1)) %>%
  mutate(Child_Sex.T1_rec = 
           recode_factor(Child_Sex.T1,
                "1" = "Male",
                "2"= "Female")
  )
```

# reading in pds t1 through t3
```{r, echo=FALSE}

pdst1 <-"~/Box/pds.t1.sav"
pdst1_df <- read_sav(pdst1) %>%
  mutate(
    ELS_ID = as.factor(ELS_ID)
    ) %>%
  left_join(.,
            demo_t1_clean,
            by = "ELS_ID")
# pds t1 includes the tk1s

pdst2 <- "~/Box/pds.T2.sav"
pdst2_df <- read_sav(pdst2) %>%
  mutate(
    ELS_ID = as.factor(ELS_ID)
    ) %>%
  left_join(.,
            demo_t1_clean,
            by = "ELS_ID")


pdst3 <- "~/Box/pds.T3.sav"
pdst3_df <- read_sav(pdst3) %>%
  mutate(
    ELS_ID = as.factor(ELS_ID)
    ) %>%
  left_join(.,
            demo_t1_clean,
            by = "ELS_ID")

```


### defining functions
```{r}
to_fact <-
  function(x) as.factor(x)
to_num <-
  function(x) as.numeric(x)
```

# Cleaning pds T1

## converting variables
```{r}
# replace with NA

pdst1_df_nareplace <- 
  pdst1_df %>%
  replace_with_na(.,
                  replace = list(pds_1_devo.T1 = c("888", "999", "10"),
                                 pds_2_growth.T1 = c("888", "999", "10"),
                                 pds_3_bodyhair.T1 = c("888", "999", "10"),
                                 pds_4_skin.T1 = c("888", "999", "10"),
                                 mpds_5_voice.T1 = c("888", "999", "10"),
                                 mpds_6_hairface.T1 = c("888", "999", "10"),
                                 fpds_5_breast.T1 = c("888", "999", "10"),
                                 fpds_menstr.T1 = c("888", "999", "10"),
                                 fpds_menst_mm.T1= c("888", "999", "10"),
                                 fpds_menst_yy.T1= c("888", "999", "10")))
pdst1_df_clean <- 
  pdst1_df_nareplace %>%
  mutate(
    ELS_ID = as.factor(ELS_ID)
  ) %>%
  dplyr::select(-Timepoint, -Session_Type)
```

## computing new variables for T1

_creating gonad, adrenal subscale, and total score for boys and girls spearately_
```{r}
pdst1_df_clean <- 
  pdst1_df_clean %>%
  mutate(pds_female_gonad.T1 = 
           ifelse(
             Child_Sex.T1_rec == "Female",
           rowMeans(
             pdst1_df[,c(
               "pds_2_growth.T1",
               "fpds_5_breast.T1",
               "fpds_menstr.T1"
               )], 
             na.rm = TRUE),
           "NA"
           )
         ) %>%
  mutate(pds_male_gonad.T1 =
           ifelse(
             Child_Sex.T1_rec == "Male",
           rowMeans(
             pdst1_df[,c(
               "pds_2_growth.T1",
               "mpds_5_voice.T1",
               "mpds_6_hairface.T1"
               )],
             na.rm = TRUE),
           "NA"
           )
         ) %>%
  mutate(pds_female_adren.T1 =
           ifelse(
             Child_Sex.T1_rec == "Female",
           rowMeans(
             pdst1_df[,c(
               "pds_3_bodyhair.T1",
               "pds_4_skin.T1"
             )],
             na.rm = TRUE),
           "NA"
           )
         ) %>%
  mutate(pds_male_adren.T1 =
           ifelse(
             Child_Sex.T1_rec == "Male",
           rowMeans(
             pdst1_df[,c(
               "pds_3_bodyhair.T1",
               "pds_4_skin.T1"               
             )],
             na.rm = TRUE),
           "NA"
           )
         )

pdst1_df_clean <-
  pdst1_df_clean %>%
  mutate_at(
    vars(
      starts_with("pds"),
      -pds_date.T1),
    to_num
  )

pdst1_df_clean$pds_female_total.T1 <- rowSums(pdst1_df_clean[, c("pds_female_gonad.T1","pds_female_adren.T1")])
pdst1_df_clean$pds_male_total.T1 <- rowSums(pdst1_df_clean[, c("pds_male_gonad.T1","pds_male_adren.T1")])

# replace with NA
pdst1_df_clean <- 
  pdst1_df_clean %>%
  replace_with_na(replace =
                    list(pds_female_total.T1 = c("NA"))) %>%
  replace_with_na(replace = 
                    list(pds_male_total.T1 = c("NA"))) %>%
  replace_with_na(replace = 
                    list(pds_female_gonad.T1 = c("NA"))) %>%
  replace_with_na(replace = 
                    list(pds_female_adren.T1 = c("NA"))) %>%
  replace_with_na(replace = 
                    list(pds_male_gonad.T1 = c("NA"))) %>%
  replace_with_na(replace = 
                    list(pds_male_adren.T1 = c("NA")))
```

```{r}
# tidying pds at t1 up
pdst1_df_tidy <-
  pdst1_df_clean %>%
  gather(pds_total.T1,
         pds_total_value.T1,
         pds_female_total.T1,
         pds_male_total.T1) %>%
  filter(!is.na(pds_total_value.T1)
         )  %>%
  gather(pds_adren.T1,
         pds_adren_value.T1,
         pds_male_adren.T1,
         pds_female_adren.T1) %>%
  filter(!is.na(pds_adren_value.T1)
         ) %>%
  gather(pds_gonad.T1,
         pds_gonad_value.T1,
         pds_male_gonad.T1,
         pds_female_gonad.T1) %>%
  filter(!is.na(pds_gonad_value.T1)
         )
```


```{r}
dup <-
  pdst1_df_tidy %>%
filter(
  duplicated(ELS_ID)
)


```

```{r}
# keeping uniques IDs
pdst1_df_tidy <-
  pdst1_df_tidy %>%
    distinct(ELS_ID,
             .keep_all = TRUE)
```


# Cleaning pds T2  

### filtering out duplicated IDs in pdst2
_I'm using the .5/.2 sessions bc they are closest to the scan date_
```{r, warning=FALSE, echo=FALSE}

pdst2_df_dup <- 
  pdst2_df %>%
  filter(duplicated(ELS_ID)
         ) 

# adding a "re-assessed" variable
pdst2_df_clean <- 
  pdst2_df %>%
  mutate(reassessed =
           ifelse(is.na(Session_Type),
                  "no",
                  "yes"
                  )
         )
```

```{r}

pdst2_df_unique <- 
  pdst2_df_clean %>%
  filter(!(ELS_ID =="xx" &
             is.na(Session_Type)
           )
  )

# keeping uniques IDs
pdst2_df_tidy <-
  pdst2_df_unique %>%
    distinct(ELS_ID,
             .keep_all = TRUE)
```


_creating gonad, adrenal subscale, and total score for boys and girls spearately_
```{r}
# replace with NA

pdst2_df_nareplace <- 
  pdst2_df_tidy %>%
  replace_with_na(.,
                  replace = list(pds_1_devo.T2 = c("888", "999", "10"),
                                 pds_2_growth.T2 = c("888", "999", "10"),
                                 pds_3_bodyhair.T2 = c("888", "999", "10"),
                                 pds_4_skin.T2 = c("888", "999", "10"),
                                 mpds_5_voice.T2 = c("888", "999", "10"),
                                 mpds_6_hairface.T2 = c("888", "999", "10"),
                                 fpds_5_breast.T2 = c("888", "999", "10"),
                                 fpds_menstr.T2 = c("888", "999", "10"),
                                 fpds_menst_mm.T2= c("888", "999", "10"),
                                 fpds_menst_yy.T2= c("888", "999", "10")))

pdst2_df_clean <- 
  pdst2_df_tidy %>%
  mutate(pds_female_gonad.T2 = 
           ifelse(
             Child_Sex.T1_rec == "Female",
           rowMeans(
             pdst2_df_tidy[,c(
               "pds_2_growth.T2",
               "fpds_5_breast.T2",
               "fpds_menstr.T2"
               )], 
             na.rm = TRUE),
           "NA"
           )
         ) %>%
  mutate(pds_male_gonad.T2 =
           ifelse(
             Child_Sex.T1_rec == "Male",
           rowMeans(
             pdst2_df_tidy[,c(
               "pds_2_growth.T2",
               "mpds_5_voice.T2",
               "mpds_6_hairface.T2"
               )],
             na.rm = TRUE),
           "NA"
           )
         ) %>%
  mutate(pds_female_adren.T2 =
           ifelse(
             Child_Sex.T1_rec == "Female",
           rowMeans(
             pdst2_df_tidy[,c(
               "pds_3_bodyhair.T2",
               "pds_4_skin.T2"
             )],
             na.rm = TRUE),
           "NA"
           )
         ) %>%
  mutate(pds_male_adren.T2 =
           ifelse(
             Child_Sex.T1_rec == "Male",
           rowMeans(
             pdst2_df_tidy[,c(
               "pds_3_bodyhair.T2",
               "pds_4_skin.T2"               
             )],
             na.rm = TRUE),
           "NA"
           )
         )

pdst2_df_clean <-
  pdst2_df_clean %>%
  mutate_at(
    vars(
      starts_with("pds"),
      -pds_date.T2),
    to_num)

pdst2_df_clean$pds_female_total.T2 <- rowSums(pdst2_df_clean[, c("pds_female_gonad.T2","pds_female_adren.T2")])
pdst2_df_clean$pds_male_total.T2 <- rowSums(pdst2_df_clean[, c("pds_male_gonad.T2","pds_male_adren.T2")])

# replace with NA
pdst2_df_clean <- 
  pdst2_df_clean %>%
  replace_with_na(replace =
                    list(pds_female_total.T2 = c("NA", "NaN"))) %>%
  replace_with_na(replace = 
                    list(pds_male_total.T2 = c("NA", "NaN"))) %>%
  replace_with_na(replace = 
                    list(pds_female_gonad.T2 = c("NA", "NaN"))) %>%
  replace_with_na(replace = 
                    list(pds_female_adren.T2 = c("NA", "NaN"))) %>%
  replace_with_na(replace = 
                    list(pds_male_gonad.T2 = c("NA", "NaN"))) %>%
  replace_with_na(replace = 
                    list(pds_male_adren.T2 = c("NA", "NaN")))


```

```{r}
# tidying pds at t2 up
pdst2_df_tidy <-
  pdst2_df_clean %>%
  gather(pds_total.T2,
         pds_total_value.T2,
         pds_female_total.T2,
         pds_male_total.T2) %>%
  filter(!is.na(pds_total_value.T2)
         )  %>%
  gather(pds_adren.T2,
         pds_adren_value.T2,
         pds_male_adren.T2,
         pds_female_adren.T2) %>%
  filter(!is.na(pds_adren_value.T2)
         ) %>%
  gather(pds_gonad.T2,
         pds_gonad_value.T2,
         pds_male_gonad.T2,
         pds_female_gonad.T2) %>%
  filter(!is.na(pds_gonad_value.T2)
         )
```


# Cleaning pds T3  
```{r}
# replace with NA

pdst3_df_nareplace <- 
  pdst3_df %>%
  replace_with_na(.,
                  replace = list(pds_1_devo.T3 = c("888", "999", "10"),
                                 pds_2_growth.T3 = c("888", "999", "10"),
                                 pds_3_bodyhair.T3 = c("888", "999", "10"),
                                 pds_4_skin.T3 = c("888", "999", "10"),
                                 mpds_5_voice.T3 = c("888", "999", "10"),
                                 mpds_6_hairface.T3 = c("888", "999", "10"),
                                 fpds_5_breast.T3 = c("888", "999", "10"),
                                 fpds_menstr.T3 = c("888", "999", "10"),
                                 fpds_menst_mm.T3= c("888", "999", "10"),
                                 fpds_menst_yy.T3= c("888", "999", "10")))
pdst3_df_clean <- 
  pdst3_df_nareplace %>%
  mutate(
    ELS_ID = as.factor(ELS_ID)
  ) %>%
  mutate_at(
    vars(
      starts_with("pds"),
      -pds_date.T3),
    to_num) %>%
  dplyr::select(-Timepoint)
```

_I'm using the .5/.2 sessions bc they are closest to the scan date_
```{r, warning=FALSE, echo=FALSE}

# duplicated IDs
pdst3_df_dup <- 
  pdst3_df_clean %>%
  filter(duplicated(ELS_ID)
         ) 

# adding a "re-assessed" variable
pdst3_df_clean <- 
  pdst3_df_clean %>%
  mutate(reassessed =
           ifelse(is.na(Session_Type),
                  "no",
                  "yes"
                  )
         )
```

```{r}

pdst3_df_unique <- 
  pdst3_df_clean %>%
  filter(!(ELS_ID == "xx" & 
           is.na(Session_Type)
           )
         )
        
# duplicated IDs
pdst3_df_dup <- 
  pdst3_df_unique  %>%
  filter(duplicated(ELS_ID)
         ) 

# keeping uniques IDs
pdst3_df_tidy <-
  pdst3_df_unique %>%
    distinct(ELS_ID,
             .keep_all = TRUE)
```

_creating gonad, adrenal subscale, and total score for boys and girls spearately_
```{r}

pdst3_df_clean <- 
  pdst3_df_tidy %>%
  mutate(pds_female_gonad.T3 = 
           ifelse(
             Child_Sex.T1_rec == "Female",
           rowMeans(
             pdst3_df_tidy[,c(
               "pds_2_growth.T3",
               "fpds_5_breast.T3",
               "fpds_menstr.T3"
               )], 
             na.rm = TRUE),
           "NA"
           )
         ) %>%
  mutate(pds_male_gonad.T3 =
           ifelse(
             Child_Sex.T1_rec == "Male",
           rowMeans(
             pdst3_df_tidy[,c(
               "pds_2_growth.T3",
               "mpds_5_voice.T3",
               "mpds_6_hairface.T3"
               )],
             na.rm = TRUE),
           "NA"
           )
         ) %>%
  mutate(pds_female_adren.T3 =
           ifelse(
             Child_Sex.T1_rec == "Female",
           rowMeans(
             pdst3_df_tidy[,c(
               "pds_3_bodyhair.T3",
               "pds_4_skin.T3"
             )],
             na.rm = TRUE),
           "NA"
           )
         ) %>%
  mutate(pds_male_adren.T3 =
           ifelse(
             Child_Sex.T1_rec == "Male",
           rowMeans(
             pdst3_df_tidy[,c(
               "pds_3_bodyhair.T3",
               "pds_4_skin.T3"               
             )],
             na.rm = TRUE),
           "NA"
           )
         )

pdst3_df_clean <-
  pdst3_df_clean %>%
  mutate_at(
    vars(
      starts_with("pds"),
      -pds_date.T3),
    to_num)

pdst3_df_clean$pds_female_total.T3 <- rowSums(pdst3_df_clean[, c("pds_female_gonad.T3","pds_female_adren.T3")])
pdst3_df_clean$pds_male_total.T3 <- rowSums(pdst3_df_clean[, c("pds_male_gonad.T3","pds_male_adren.T3")])

# replace with NA
pdst3_df_clean <- 
  pdst3_df_clean %>%
  replace_with_na(replace =
                    list(pds_female_total.T3 = c("NA", "NaN"))) %>%
  replace_with_na(replace = 
                    list(pds_male_total.T3 = c("NA", "NaN"))) %>%
  replace_with_na(replace = 
                    list(pds_female_gonad.T3 = c("NA", "NaN"))) %>%
  replace_with_na(replace = 
                    list(pds_female_adren.T3 = c("NA", "NaN"))) %>%
replace_with_na(replace = 
                    list(pds_male_gonad.T3 = c("NA", "NaN"))) %>%
  replace_with_na(replace = 
                    list(pds_male_adren.T3 = c("NA", "NaN")))


```

```{r}
# tidying pds at t3 up
pdst3_df_tidy <-
  pdst3_df_clean %>%
  gather(pds_total.T3,
         pds_total_value.T3,
         pds_female_total.T3,
         pds_male_total.T3) %>%
  filter(!is.na(pds_total_value.T3)
         )  %>%
  gather(pds_adren.T3,
         pds_adren_value.T3,
         pds_male_adren.T3,
         pds_female_adren.T3) %>%
  filter(!is.na(pds_adren_value.T3)
         ) %>%
  gather(pds_gonad.T3,
         pds_gonad_value.T3,
         pds_male_gonad.T3,
         pds_female_gonad.T3) %>%
  filter(!is.na(pds_gonad_value.T3)
         )
```

