---
title: "1_tanner_threetimpts_cleaning_script"
author: "Jackie"
date: "2/11/2020"
output: html_notebook 
---

# loading libraries
```{r}
library(lmerTest)
library(tidyverse)
library(foreign)
library(haven)
library(readxl)
library(naniar)
```

# read in tanner info

## reading in tanner t1 through t3
```{r, echo=FALSE}
tannert1 <-"~/Box/tanner.T1.sav"
tannert1_df <- read_sav(tannert1)

tannertk1 <-  "~/Box/demographics_golden_standard.TK1.sav"
tannertk1_df <- read_sav(tannertk1)

tannert2 <- "~/Box/tanner.T2.sav"
tannert2_df <- read_sav(tannert2)

tannert3 <- "~/Boxtanner.T3.sav"
tannert3_df <- read_sav(tannert3)
```

### defining functions
```{r}
to_fact <-
  function(x) as.factor(x)
to_num <-
  function(x) as.numeric(x)
```

# Cleaning Tanner T1

## converting variables
```{r}
tannert1_df_clean <- 
  tannert1_df %>%
  mutate(
    ELS_ID = as.factor(ELS_ID)
  ) %>%
  mutate_at(
    vars(
      starts_with("tanner"),
      -tanner_date.T1),
    to_num) %>%
  dplyr::select(-Timepoint)
  
```

## computing new variables for T1

_averaged breast/pub for female and averaged testes/pub for male_
```{r}
tannert1_df_clean <- 
  tannert1_df_clean %>%
  mutate(tanner_female_avg.T1 = 
           rowMeans(
             tannert1_df_clean[,c(
               "tanner_female_breasts.T1",
               "tanner_female_pubhair.T1"
               )], 
             na.rm = T)
         ) %>%
  mutate(tanner_male_avg.T1 =
           rowMeans(
             tannert1_df_clean[,c(
               "tanner_male_testes.T1",
               "tanner_male_pubhair.T1"
               )],
             na.rm = T)
         )

# replace with NA
tannert1_df_clean <- 
  tannert1_df_clean %>%
  replace_with_na(replace =
                    list(tanner_female_avg.T1 = c("NaN"))) %>%
  replace_with_na(replace = 
                    list(tanner_male_avg.T1 = c("NaN"))) 
```

### cleaning up NaNs and excluding those who came in over tanner stage 3
```{r}
tannert1_df_clean <- 
  tannert1_df_clean %>%
  mutate(above_tanner_limit =
           ifelse(tanner_female_avg.T1 > 3 |
                    tanner_male_avg.T1 > 3,
                  "1",
                  "0")
         )
  # mutate(above_tanner_limit =
  #          fct_explicit_na(above_tanner_limit,
  #                          na_level = "NA")
  #        )
```

## removing subjects above tanner stage 3
```{r}
# not removing for now
above_tanner_3 <-
  tannert1_df_clean %>%
  group_by(above_tanner_limit) %>%
  summarize(N = n()
            )
above_tanner_3 <-
  tannert1_df_clean %>%
  filter(above_tanner_limit == "1")
# 10 subjects (F = 6, M = 4) above tanner stage 3 upon entering study, so
# will remove these subjects
tannert1_df_clean_noNA <-
  tannert1_df_clean %>%
  filter(is.na(above_tanner_limit)) %>%
  select(-above_tanner_limit)
```

```{r}
# tidying tanner at t1 up
tannert1_df_tidy <-
  tannert1_df_clean %>%
  gather(tannert1,
         tanner_average.T1,
         tanner_male_avg.T1,
         tanner_female_avg.T1) %>%
  filter(!is.na(tanner_average.T1)
         ) 
```


```{r}
dup <-
  tannert1_df_tidy %>%
filter(
  !is.na(Session_Type)
)
```


```{r}
tannert1_df_clean_filtered <-
  tannert1_df_tidy %>%
  filter(!(ELS_ID == "xx" &
           tanner_date.T1 == "yy")
  )
```


#### merging with tk1
```{r}
# tanner t1 with tanner tk1
tannertk1_df <-
  tannertk1_df %>%
  mutate(ELS_ID = as.factor(ELS_ID),
         TK1_Child_Sex_R = as.factor(TK1_Child_Sex_R))

xtra_tk1info <- "~/TK1_Demographics_Tanner.sav"
xtra_tk1 <- read_sav(xtra_tk1info) %>%
  mutate(ELS_ID = as.factor(ELS_ID))

tannertk1 <-
  left_join(
    tannertk1_df,
    xtra_tk1,
    by = "ELS_ID"
  ) %>%
  mutate(ELS_ID = as.factor(ELS_ID))

tannert1_df_clean_filtered <- 
  left_join(
    tannert1_df_clean_filtered,
    tannertk1,
    by = "ELS_ID")

tannert1_df_clean_filtered <-
  tannert1_df_clean_filtered %>%
  mutate(
    ELS_ID = as.factor(ELS_ID),
    tanner_date.T1 =
      ifelse(
        is.na(tanner_date.T1)  &
          !is.na(Session_Date),
        Session_Date,
        tanner_date.T1
      ),
    tanner_date.T1 = as.Date(tanner_date.T1,
                             origin = "1970-01-01"),
    Age.TK1 = Age_at_session,
    tanner_female_avg.T1 =
      ifelse(
        is.na(tanner_average.T1) &
          !is.na(TK1_TannerAvg),
        TK1_TannerAvg,
        tanner_average.T1
      )
  ) 
```


### writing T1 tanner file
```{r}
tanner_t1_fp <- 
  "~/tanner_t1.csv"
write_csv(tannert1_df_clean_filtered,tanner_t1_fp)
```

# Cleaning Tanner T2  
__Includes tanner t2 s1 and tanner t2 s2__

## converting variables
```{r}
tannert2_df_clean <- 
  tannert2_df %>%
  mutate(ELS_ID = as.factor(ELS_ID)
  ) %>%
  mutate_at(
    vars(
      starts_with("tanner"),
      -tanner_date_s1.T2,
      -tanner_date_s2.T2),
    to_num
  )
```

### filtering out duplicated IDs in tannert2
_I'm using the .5/.2 sessions bc they are closest to the scan date_
```{r, warning=FALSE, echo=FALSE}

# duplicated IDs
tannert2_df_dup <- 
  tannert2_df_clean  %>%
  filter(duplicated(ELS_ID)
         ) 

# adding a "re-assessed" variable
tannert2_df_clean <- 
  tannert2_df_clean %>%
  mutate(reassessed =
           ifelse(is.na(Session_Type),
                  "no",
                  "yes"
                  )
         )
```

```{r}
tannert2_df_unique <- 
  tannert2_df_clean %>%
  filter(!(ELS_ID == "xx" & 
           is.na(Session_Type)
           )
         ) 
```


## creating variable to denote assessed at scan
_t2s2 tanner was recorded for those who had tanner reassesed at their scan._  
_A majority of subjects who are in both t2s1 and t2s2 tanner data files_   
_had there scan closest to the t2s2 tanner date._  
```{r}
tannert2_note_scan <-
  tannert2_df_unique %>%
  mutate(tanner_at_scan =
           1) %>%
  mutate(tanner_at_scan =
           ifelse(ELS_ID == "x1"| 
                    ELS_ID == "x2" |
                    ELS_ID == "x3"|
                    ELS_ID == "x4" |
                    ELS_ID == "x5",
                  0,
                  tanner_at_scan)
  ) # 1 = tanner assessed at scan; 0 = no scan
```


## cleaning the t2final file
```{r}
tannert2_final <- 
  tannert2_note_scan %>%
  mutate(has_t2s2 =
           ifelse(
             !is.na(tanner_average_s2.T2),
             1,
             0)
  )

tannert2_final_clean <- 
  tannert2_final %>%
  mutate(tanner_average.T2 =
           ifelse(
             has_t2s2 == "0",
             tanner_average_s1.T2,
             tanner_average_s2.T2
           )
         ) %>%
  mutate(tanner_date.T2 =
           ifelse(
             has_t2s2 == "0",
             tanner_date_s1.T2,
             tanner_date_s2.T2)
         ) %>%
  mutate(tanner_date.T2 = 
           as.Date(tanner_date.T2,
                   origin = "1970-01-01")
         )
```

#### check of NAs
```{r}
anynas <- 
  tannert2_final_clean  %>%
  filter(
    is.na(tanner_average.T2)
         )
```

### writing T2_final tanner file
```{r}
tanner_t2_fp <- 
  "~/Box/tanner_t2.csv"
write_csv(tannert2_final_clean,tanner_t2_fp)
```

# summary 

## How many subjects at T1?
```{r}
tannert1_table <-
  tannert1_df_clean_filtered %>%
  group_by(tannert1) %>%
  summarize(
    tanner_avg = mean(tanner_average.T1, na.rm = TRUE),
    tanner_med = median(tanner_average.T1, na.rm = TRUE),
    n = n()
  )
tannert1_table
# 131 F
# 93 M
t.test(tannert1_df_clean_filtered$tanner_average.T1 ~  tannert1_df_clean_filtered$tannert1,
       na.action = na.omit)
# t = 1.4311, df = 212.53, p-value = 0.1539
# mean of x = 2.085878 mean of y = 1.946237
```

# How many subjects at T2?
```{r}
tannert2_table <- 
  tannert2_final_clean %>%
  filter(!is.na(tanner_average.T2)) %>%
  summarize(
    n = n()
  )
tannert2_table
```

# Cleaning Tanner T3  
```{r}
tannert3_df_clean <- 
  tannert3_df %>%
  mutate(
    ELS_ID = as.factor(ELS_ID)
  ) %>%
  mutate_at(
    vars(
      starts_with("tanner"),
      -tanner_date.T3),
    to_num) %>%
  dplyr::select(-Timepoint)
```

_I'm using the .5/.2 sessions bc they are closest to the scan date_
```{r, warning=FALSE, echo=FALSE}

tannert3_df_dup <- 
  tannert3_df_clean  %>%
  filter(duplicated(ELS_ID)
         ) 

# adding a "re-assessed" variable
tannert3_df_clean <- 
  tannert3_df_clean %>%
  mutate(reassessed =
           ifelse(is.na(Session_Type),
                  "no",
                  "yes"
                  )
         )
```

```{r}
tannert3_df_unique <- 
  tannert3_df_clean %>%
  filter(!(ELS_ID == "xx" & 
           is.na(Session_Type)
           )
         ) 
```

# cleaning T3 tanner average variable name
```{r}
# 1 = male (converting to "0")
# 2 = female (converting to "1")
tannert3_final_clean <-
  tannert3_df_unique %>%
  mutate(
    tanner_sex.T3 = as.factor(tanner_sex.T3)
  ) %>%
  rename(
    tanner_average.T3 = ends_with("average.T3")
  ) %>%
  mutate(
    recode_factor(tanner_sex.T3,
                  "1" = "0",
                  "2" = "1")
  )
```

#### check of NAs
```{r}
anynas <- 
  tannert3_final_clean %>%
  filter(
    is.na(tanner_average.T3)
         )
```

# summary 

## How many subjects at T3?
```{r}
tannert3_table <-
  tannert3_final_clean %>%
  group_by(tanner_sex.T3) %>%
  summarize(
    tanner_avg = mean(tanner_average.T3, na.rm = TRUE),
    tanner_med = median(tanner_average.T3, na.rm = TRUE),
    n = n()
  )
tannert3_table
# 69 F
# 59 M
t.test(tannert3_final_clean$tanner_average.T3 ~  tannert3_final_clean$tanner_sex.T3,
       na.action = na.omit)
# t = 0.16893, df = 117.75, p-value = 0.8661
# mean of x = 4.296610 mean of y = 4.275362
```


### writing T3_final tanner file
```{r}
tanner_t3_fp <- 
  "~/Box/tanner_t3.csv"
write_csv(tannert3_final_clean,tanner_t3_fp)
```
