---
title: "2_mw_sleep"
author: "Jackie"
date: "2/10/2020"
output: html_notebook
---
# loading libraries
```{r libraries}
library(tidyverse)
library(foreign)
library(haven)
library(readxl)
library(lubridate)
```


# Read in MW sleep data
```{r read mw sleep}
mw_sleep_fp <- "~/Box/Daily Sleep 6_45am Survey-20200214144154.csv"
mw_sleep <- read_csv(mw_sleep_fp)
```

# cleaning MW_daily_sleep
```{r cleaning mw sleep}
mw_sleep_clean <-
  mw_sleep %>%
  rename(
    User_ID = 
      `User Id`,
    ELS_ID =
      `First Name`,
    triggerdate =
      `Trigger Date`,
    DeviceOS = 
      `Device OS`,
    DailySleep_hrs = 
      `How many hours did you sleep last night?`,
    DailySleep_satisfaction = 
      `How restful or satisfying was your sleep last night?`
  ) %>%
  mutate(
    User_ID = as.factor(User_ID)
  ) %>%
  dplyr::select(-c(`Response Id`,
            `Submission Location`,
            `Trigger Id`,
            `Device Id`)
  ) %>%
  mutate(
    DailySleep_hrs = as.factor(DailySleep_hrs)
      ) %>%
  mutate(
    DailySleep_hrs_rec =
      recode_factor(DailySleep_hrs,
                    "5hr or less" = 5,
                    "6 hr." = 6,
                    "7hr." = 7,
                    "8hr." = 8,
                    "9hr. or more" = 9)
  ) %>%
  mutate(
    DailySleep_hrs_rec = 
      as.character(DailySleep_hrs_rec)
         ) %>%
  mutate(DailySleep_hrs_rec =
           as.numeric(DailySleep_hrs_rec)
         )
```


## pulling out only "c"
```{r only child sleep}
mw_sleep_c <-
  mw_sleep_clean %>%
  mutate(
    porc =
      ifelse(str_detect(ELS_ID, "c"),
             "C",
             "P"
             )
    )

mw_sleep_c <-
  mw_sleep_c %>%
  filter(
    porc == "C"
  )

mw_sleep_c_clean <-
  mw_sleep_c %>%
  mutate(
    ELS_ID =
      str_remove_all(ELS_ID,
                     "[c]")
    ) %>%
  mutate(ELS_ID =
           as.numeric(ELS_ID)
         ) %>%
  dplyr::select(
    ELS_ID,
    triggerdate,
    DailySleep_hrs,
    DailySleep_satisfaction,
    DailySleep_hrs_rec
  ) %>%
  mutate(
    ELS_ID =
      as.factor(ELS_ID)
  )
```

### parsing dates
```{r parsing dates}

mw_sleep_c_clean <-
  mw_sleep_c_clean %>%
  mutate(
    date1 = parse_date(triggerdate, "%d/%m/%y"),
    date2 = parse_date(triggerdate, "%d/%m/%Y"),
    date1 = ymd(date1),
    date2 = ymd(date2),
    date1 = as.character(date1),
    date2 = as.character(date2)
  )
```

## creating day and week variables
```{r}
# creating day and week variables
mw_sleep_c_clean_day <-
  mw_sleep_c_clean %>%
  mutate(
    date = date1,
    date = ifelse(
      !is.na(date2),
      date2,
      date
      )
    ) %>%
  mutate(
    day1 = day(date1),
    day2 = day(date2),
    wday1 = wday(date1, label = TRUE),
    wday2 = wday(date2, label = TRUE)
  ) %>%
  mutate(
    day1 = as.character(day1),
    day2 = as.character(day2),
    wday1 = as.character(wday1),
    wday2 = as.character(wday2)
  ) %>%
  mutate(
    day = day1,
    day = ifelse(
      !is.na(day2),
      day2,
      day
    ),
    week = wday1,
    week = ifelse(
      !is.na(wday2),
      wday2,
      week
      )
    ) %>%
  dplyr::select(
    -c(day1:wday2)
  )

# day num
mw_sleep_c_clean_dayorder  <-
mw_sleep_c_clean_day %>%
  left_join(mw_sleep_c_clean_day %>% 
                        group_by(ELS_ID) %>% 
                        mutate(dayorder=order(date))) %>%
  mutate(
    week = as.factor(week)
  ) %>%
  ungroup()
```

## wkday or wkend
```{r}
mw_sleep_c_clean_dayorder <-
  mw_sleep_c_clean_dayorder %>%
  mutate(
    wkday =
      ifelse(
        week == "Sun" |
          week == "Sat",
        "wkend",
        "wkday"
      )
  )
```


# dealing with not enough or missing data  

if there are fewer than 2 days of data, I'm removing the subject
```{r not enough data}
# how many days per individual?
mw_sleep_c_clean_dayorder <-
  mw_sleep_c_clean_dayorder %>%
  group_by(ELS_ID) %>%
  mutate(
    data_num_obs =
      max(dayorder)
  ) %>%
  ungroup()

# filter those who have more than 1 day
mw_sleep_filter  <-
  mw_sleep_c_clean_dayorder %>%
  group_by(ELS_ID) %>%
  filter(
    data_num_obs > 1
  ) %>%
  dplyr::select(
    -c(
    triggerdate, date1, date2, day
    )
    ) %>%
  ungroup()
```

# data viz
```{r sleep sat}
mw_sleep_filter %>%
  ggplot(aes(dayorder, DailySleep_satisfaction,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r sleep hrs}
mw_sleep_filter %>%
  ggplot(aes(dayorder, DailySleep_hrs_rec,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

### distribution of weekday vs. weekend and association with sleep sat and hrs
```{r}
mw_sleep_filter %>%
  ggplot(
    aes(x = week, y = DailySleep_satisfaction )
  ) +
  geom_boxplot()

mw_sleep_filter %>%
  ggplot(
    aes(x = wkday, y = DailySleep_satisfaction )
  ) +
  geom_boxplot()

mw_sleep_filter %>%
  ggplot(
    aes(x = week, y = DailySleep_hrs_rec )
  ) +
  geom_boxplot()
mw_sleep_filter %>%
  ggplot(
    aes(x = wkday, y = DailySleep_hrs_rec )
  ) +
  geom_boxplot()

```

### write
```{r}
write_csv(mw_sleep_filter, "~/Box/MW_daily_sleep.csv")
```

### if I want reduced sleep data
```{r}
mw_sleep_red <-
  mw_sleep_filter %>%
  group_by(ELS_ID, wkday) %>%
  summarize(
    dailysleep_sat_mean = mean(DailySleep_satisfaction, na.rm = TRUE),
    dailysleep_sat_sd = sd(DailySleep_satisfaction, na.rm = TRUE),
    dailysleep_hrs_mean = mean(DailySleep_hrs_rec, na.rm = TRUE),
    dailysleep_hrs_sd = sd(DailySleep_hrs_rec, na.rm = TRUE)
  )
```

```{r}
summary(lm(dailysleep_sat_mean ~ wkday, data = mw_sleep_red))
#  no sig diff btwn sleep sat on wkday vs. wkend
```

### write
```{r}
write_csv(mw_sleep_red, "~/Box/mw_sleep_red.csv")
```

