---
title: "2_cleaning_mw_mood"
author: "Jackie S. Kirsh"
date: "02/10/2020"
output: html_notebook
---


# loading libraries
```{r}
library(tidyverse)
library(foreign)
library(haven)
library(readxl)
```

# Read in MW mood data
```{r}
mw_mood_fp <- "~/Box/Daily Emotion-20200214144103.csv"
mw_mood <- read_csv(mw_mood_fp)
```


# cleaning MW daily emo
```{r}
mw_mood_clean <-
  mw_mood %>%
  rename(
    User_ID = 
      `User Id`,
    ELS_ID =
      `First Name`,
    DeviceOS = 
      `Device OS`,
    DailyEMO_sad = 
      `Since the last prompt, please indicate the extent to which you felt sad`,
    DailyEMO_annoyedorangry = 
      `Since the last prompt, please indicate the extent to which you felt annoyed/angry`,
    DailyEMO_grouchyorcranky =
      `Since the last prompt, please indicate the extent to which you felt grouchy/cranky`,
    DailyEMO_worried =
      `Since the last prompt, please indicate the extent to which you felt worried`,
    DailyEMO_anxious =
      `Since the last prompt, please indicate the extent to which you felt anxious`,
    DailyEMO_happy =
      `Since the last prompt, please indicate the extent to which you felt happy`,
    DailyEMO_cheerful =
      `Since the last prompt, please indicate the extent to which you felt cheerful`,
    DailyEMO_excited =
    `Since the last prompt, please indicate the extent to which you felt excited`,
    DailyEMO_energetic =
      `Since the last prompt, please indicate the extent to which you felt energetic`,
    DailyEMO_bored =
      `Since the last prompt, please indicate the extent to which you felt bored`,
    DailyEMO_want_other_people =
      `Since the last prompt, please indicate the extent to which you want to be with other people`,
    DailyEMO_want_alone =
      `Since the last prompt, please indicate the extent to which you want to be alone`,
    DailyEMO_who_are_you_with =
      `Right now, who are you with?`,
    DailyEMO_facetoface_YesorNo =
      `Right now, are you having a face-to-face conversation?`,
    DailyEMO_digital_convo_YesorNo =
      `Right now, are you having a real-time digital (phone, text, Facebook, video) conversation?`
  ) %>%
  mutate(
    User_ID = as.factor(User_ID)
  ) %>%
  select(-`Response Id`)
```

# pulling out only "c"
```{r}
mw_mood_c <-
  mw_mood_clean %>%
  mutate(
    porc =
      ifelse(str_detect(ELS_ID, "c"),
             "C",
             "P"
             )
    )

mw_mood_c <-
  mw_mood_c %>%
  filter(
    porc == "C"
  )

mw_mood_c_clean <-
  mw_mood_c %>%
  mutate(
    ELS_ID =
      str_remove_all(ELS_ID,
                     "[c]")
    ) %>%
  mutate(ELS_ID =
           as.numeric(ELS_ID)
         ) %>%
  select(
    ELS_ID,
    `Trigger Date`,
    `Trigger Time`,
    starts_with("Daily")
  ) %>%
  mutate(
    ELS_ID =
      as.factor(ELS_ID)
  )
```

# For collapsed mood

## means and sds of variables
```{r}
mw_mood_c_collapse <-
  mw_mood_c_clean %>% 
  group_by(ELS_ID) %>%
   summarise(DailyEMO_sad_mean = mean(DailyEMO_sad, na.rm=TRUE),  
             DailyEMO_sad_sd = sd(DailyEMO_sad, na.rm=TRUE), 
             DailyEMO_annoyorang_mean = 
               mean(DailyEMO_annoyedorangry, na.rm=TRUE), 
             DailyEMO_annoyorang_sd =
               sd(DailyEMO_annoyedorangry, na.rm = TRUE),
             DailyEMO_grouchorcrank_mean = 
               mean(DailyEMO_grouchyorcranky, na.rm=TRUE),
             DailyEMO_grouchorcrank_sd = 
               sd(DailyEMO_grouchyorcranky, na.rm=TRUE),
             DailyEMO_worried_mean =
               mean(DailyEMO_worried, na.rm = TRUE),
             DailyEMO_worried_sd =
               sd(DailyEMO_worried, na.rm = TRUE),
             DailyEMO_anx_mean =
               mean(DailyEMO_anxious, na.rm = TRUE),
             DailyEMO_anx_sd =
               sd(DailyEMO_anxious, na.rm = TRUE),
             DailyEMO_happy_mean = 
               mean(DailyEMO_happy, na.rm=TRUE), 
             DailyEMO_happy_sd = 
               sd(DailyEMO_happy, na.rm=TRUE), 
             DailyEMO_cheerful_mean = 
               mean(DailyEMO_cheerful, na.rm=TRUE), 
             DailyEMO_cheerful_sd = 
               sd(DailyEMO_cheerful, na.rm=TRUE), 
             DailyEMO_excited_mean = 
               mean(DailyEMO_excited, na.rm=TRUE), 
             DailyEMO_excited_sd = 
               sd(DailyEMO_excited, na.rm=TRUE), 
             DailyEMO_energetic_mean =
               mean(DailyEMO_energetic, na.rm=TRUE),
             DailyEMO_energetic_sd= 
               sd(DailyEMO_energetic, na.rm=TRUE), 
             DailyEMO_bored_mean =
               mean(DailyEMO_bored, na.rm=TRUE),
             DailyEMO_bored_sd =
               sd(DailyEMO_bored, na.rm=TRUE),
             DailyEMO_want_other_people_mean = 
               mean(DailyEMO_want_other_people, na.rm=TRUE), 
             DailyEMO_want_alone_mean = 
               mean(DailyEMO_want_alone, na.rm=TRUE)
             )
```


# prelim pca for mw mood data

```{r}
library(broom)
library(ggfortify)
library(factoextra)

# compute pca
mw_pca <-
  mw_mood_c_collapse %>%
  filter(
    !is.na(DailyEMO_sad_sd)
    ) %>%
  dplyr::select(-ELS_ID,
                -ends_with("sd"))
  
res.pca_mw <- prcomp(mw_pca, scale = TRUE)

# visualize eigenvalues (scree plot). 
# show percentage of variances explained by each princ comp
fviz_eig(res.pca_mw)

# graph of individuals. individual with a similar profile are grped together
fviz_pca_ind(res.pca_mw,
             col.ind = "cos2", # color by quality of representation
             gradient.cols = c("lightseagreen", "skyblue1", "snow4"),
             repel = TRUE # avoid text overlap
             )

# biplot of individuals and variables
fviz_pca_biplot(res.pca_mw, repel = TRUE,
                col.var = "sienna1",
                col.ind = "skyblue1")

# PCA Results
# eigenvalues
eig.val <- get_eigenvalue(res.pca_mw)
eig.val
# keeping eigenvalues greater than 1, which are dims 1 and 2

# Results for Variables
res.var <- get_pca_var(res.pca_mw)
res.var$coord          # Coordinates
res.var$contrib        # Contributions to the PCs
res.var$cos2           # Quality of representation 
# Results for individuals
res.ind <- get_pca_ind(res.pca_mw)
res.ind$coord          # Coordinates
res.ind$contrib        # Contributions to the PCs
res.ind$cos2           # Quality of representation 
```

We essentailly get 2 factors: combining 
__negative:__ `DailyEMO_sad_mean`, `DailyEMO_annoyorang_mean`,   
`DailyEMO_grouchorcrank_mean`, `DailyEMO_worried_mean`,  
`DailyEMO_anx_mean`, `DailyEMO_bored_mean`, `DailyEMO_want_alone_mean`

__positive:__ `DailyEMO_happy_mean`, `DailyEMO_cheerful_mean`,  
`DailyEMO_excited_mean`, `DailyEMO_energetic_mean`, `DailyEMO_want_other_people_mean`
```{r}
mw_fact <-
  mw_mood_c_collapse %>%
  filter(
    !is.na(DailyEMO_sad_sd)
    ) %>%
  dplyr::select(-ends_with("sd")) %>%
  mutate(
    neg_total =
      select(., DailyEMO_sad_mean,
                  DailyEMO_annoyorang_mean,
                  DailyEMO_grouchorcrank_mean,
                  DailyEMO_worried_mean,
                  DailyEMO_anx_mean,
                  DailyEMO_bored_mean,
                  DailyEMO_want_alone_mean) %>%
      rowSums(na.rm = TRUE),
    pos_total =
      select(., DailyEMO_happy_mean,
             DailyEMO_cheerful_mean,
             DailyEMO_excited_mean,
             DailyEMO_energetic_mean,
             DailyEMO_want_other_people_mean) %>%
      rowSums(na.rm = TRUE)
  )

```

```{r}
write_csv(mw_fact,"~/Box/mw_fact.csv")
```

