---
title: "3_tanner/pds_and_age_over_time"
author: "Jackie"
date: "2/11/2020"
output: html_notebook
---

# loading libraries
```{r}
library(lmerTest)
library(tidyverse)
library(foreign)
library(haven)
library(readxl)
library(naniar)
library(lubridate)
```


# reading in tanner and pds plus dem dataframes
This is the output from my first two scripts
```{r}
# t1

t1_fp <- "~/Box/dem_plus_tanandpds_t1.csv"
t1 <- read_csv(t1_fp) %>%
  mutate(ELS_ID = as.factor(ELS_ID))

# t2

t2_fp <- "~/Box/dem_plus_tanandpds_t2.csv"
t2 <- read_csv(t2_fp) %>%
  mutate(ELS_ID = as.factor(ELS_ID))

# t3

t3_fp <- "~/Box/dem_plus_tanandpds_t3.csv"
t3 <- read_csv(t3_fp) %>%
  mutate(ELS_ID = as.factor(ELS_ID))

```

# merging
```{r}
merged <-
  left_join(
    t1,
    t2,
    by = "ELS_ID"
  )

merged <-
  left_join(
    merged,
    t3,
    by = "ELS_ID"
  )
```

# decreasers only for tanner
literature includes decreasers if looking at pds
```{r}
merged <-
  merged %>%
  mutate(
    tanner_average.T1 = as.numeric(tanner_average.T1),
    tanner_average.T2 = as.numeric(tanner_average.T2),
    tanner_average.T3 = as.numeric(tanner_average.T3)
  ) %>%
  mutate(
    decreaser =
      ifelse(
        tanner_average.T1 > tanner_average.T2 |
          tanner_average.T1 > tanner_average.T3,
        "yes",
        "no"
      ),
    decreaser =
      ifelse(
        tanner_average.T2 > tanner_average.T3,
        "yes",
        decreaser
      )
  )
```


# exculsion?
We are excluding anyone with only one time point of data and those who "decreased"
```{r}
merged_exc <-
  merged %>%
  mutate(
    num_timepts =
      ifelse(
        is.na(pds_total_value.T1) &
          is.na(pds_total_value.T2) &
          is.na(pds_total_value.T3),
        "0",
        "1"
      ),
    num_timepts =
      ifelse(
        !is.na(pds_total_value.T1) &
          is.na(pds_total_value.T2) &
          is.na(pds_total_value.T3),
        "1",
        num_timepts
      ),
    num_timepts =
      ifelse(
        is.na(pds_total_value.T1) &
          !is.na(pds_total_value.T2) &
          is.na(pds_total_value.T3),
        "1",
        num_timepts
        ),
    num_timepts =
      ifelse(
        is.na(pds_total_value.T1) &
          is.na(pds_total_value.T2) &
          !is.na(pds_total_value.T3),
        "1",
        num_timepts
        ),
    num_timepts =
      ifelse(
        !is.na(pds_total_value.T1) &
          !is.na(pds_total_value.T2) &
          is.na(pds_total_value.T3),
        "2",
        num_timepts
      ),
      num_timepts =
      ifelse(
        !is.na(pds_total_value.T1) &
          is.na(pds_total_value.T2) &
          !is.na(pds_total_value.T3),
        "2",
        num_timepts
      ),
    num_timepts = 
      ifelse(
        is.na(pds_total_value.T1) &
          !is.na(pds_total_value.T2) &
          !is.na(pds_total_value.T3),
        "2",
        num_timepts
      ),
    num_timepts = 
      ifelse(
        !is.na(pds_total_value.T1) &
          !is.na(pds_total_value.T2) &
          !is.na(pds_total_value.T3),
        "3",
        num_timepts
        )
  )

merged_d <-
  merged_exc %>%
  filter(
    is.na(decreaser) | decreaser == "no"
  )
```

### only one time point so far
```{r}
only_one_timept <-
  merged_d %>%
  filter(
    num_timepts == "1"
  )

```

#### summary table for only one time point so far
```{r}
summary_only_one_timept <-
  only_one_timept %>%
  group_by(Child_Sex.T1_rec) %>%
  summarize(
    mean_tanner = mean(tanner_average.T1, na.rm = TRUE),
    mean_pds = mean(pds_total_value.T1, na.rm = TRUE),
    median_tanner = median(tanner_average.T1, na.rm = TRUE),
    median_pds = median(pds_total_value.T1, na.rm = TRUE),
    min_tanner = min(tanner_average.T1, na.rm = TRUE),
    min_pds = min(pds_total_value.T1, na.rm = TRUE),
    max_tanner = max(tanner_average.T1, na.rm = TRUE),
    max_pds = max(tanner_average.T1, na.rm = TRUE),
    mean_age = mean(Age.T1, na.rm = TRUE),
    median_age = median(Age.T1, na.rm = TRUE),
    min_age = min(Age.T1, na.rm = TRUE),
    max_age = max(Age.T1, na.rm = TRUE),
    n = n()
  )
print(summary_only_one_timept)
```


## those with at least 2 time points
```{r}
two_or_more_tanner_pds_timepts <-
  merged_exc %>%
  filter(
    num_timepts != "1"
  ) %>%
  mutate(
    interval_t1t2 = Age.T2 - Age.T1,
    interval_t1t3_t3bx = Age_Bx.T3 - Age.T1,
    interval_t1t3_t3sc = Age_scan.T3 - Age.T1,
    interval_t2t3_t3bx = Age_Bx.T3 - Age.T2,
    interval_t2t3_t3sc = Age_scan.T3 - Age.T1
  )
```

# wide to long
```{r}
two_or_more_tanner_pds_timepts_long <-
  two_or_more_tanner_pds_timepts %>%
  mutate(
    ELS_ID = as.factor(ELS_ID)
  ) %>%
  gather(
    pds_total_timepoint,
    pds_total_value,
    pds_total_value.T1,
    pds_total_value.T2,
    pds_total_value.T3
  ) %>%
  gather(
    pds_adren_timepoint,
    pds_adren_value,
    pds_adren_value.T1,
    pds_adren_value.T2,
    pds_adren_value.T3
  ) %>%
  gather(
    pds_gonad_timepoint,
    pds_gonad_value,
    pds_gonad_value.T1,
    pds_gonad_value.T2,
    pds_gonad_value.T3
  ) %>%
  mutate(
    match_pds = 
      ifelse(
        pds_total_timepoint == "pds_total_value.T1" &
          pds_adren_timepoint == "pds_adren_value.T1" &
          pds_gonad_timepoint == "pds_gonad_value.T1",
        "match_pds",
        "no_match_pds"
      ),
    match_pds =
      ifelse(
        pds_total_timepoint == "pds_total_value.T2" &
          pds_adren_timepoint == "pds_adren_value.T2" &
          pds_gonad_timepoint == "pds_gonad_value.T2",
        "match_pds",
        match_pds
      ),
    match_pds =
      ifelse(
        pds_total_timepoint == "pds_total_value.T3" &
          pds_adren_timepoint == "pds_adren_value.T3" &
          pds_gonad_timepoint == "pds_gonad_value.T3",
        "match_pds",
        match_pds
      )      
  ) %>%
  filter(
    match_pds == "match_pds"
  ) %>% 
  gather(
    tanner_timepoint,
    tanner_score,
    tanner_average.T1,
    tanner_average.T2,
    tanner_average.T3
  ) %>%
  gather(
    age_timepoint,
    age,
    Age.T1,
    Age.T2,
    Age.T3
  ) %>%
  mutate(
    match =
      ifelse(
        tanner_timepoint == "tanner_average.T1" &
          age_timepoint == "Age.T1" &
          pds_total_timepoint == "pds_total_value.T1",
        "match",
        "nonmatch"
      ),
    match = 
      ifelse(
        tanner_timepoint == "tanner_average.T2" &
          age_timepoint == "Age.T2" &
          pds_total_timepoint == "pds_total_value.T2",
        "match",
        match       
      ),
    match =
      ifelse(
        tanner_timepoint == "tanner_average.T3" &
          age_timepoint == "Age.T3" &
          pds_total_timepoint == "pds_total_value.T3",
        "match",
        match
        )
  ) %>%
  filter(
    match == "match"
  ) %>%
  dplyr::select(
    -starts_with("match")
  ) %>%
  mutate(
    ELS_ID = as.factor(ELS_ID)
  )
```


### writing wide format dataframe
```{r}
two_or_more_tanner_pds_timepts_wide_fp <- "~/Box/two_or_more_tanner_pds_timepts_wide.csv"
write_csv(two_or_more_tanner_pds_timepts, two_or_more_tanner_pds_timepts_wide_fp)
```

```{r}
two_or_more_tanner_pds_timepts_long_fp <- "~/Box/two_or_more_tanner_pds_timepts_long.csv"
write_csv(two_or_more_tanner_pds_timepts_long, two_or_more_tanner_pds_timepts_long_fp)
```

