---
title: "4_mw_mood_long"
author: "Jackie S. Kirsh"
date: "02/10/2020"
output: html_notebook
---


# loading libraries
```{r}
library(tidyverse)
library(foreign)
library(haven)
library(readxl)
library(lubridate)
```

# Read in MW mood data
```{r}
mw_mood_fp <- "~/Box/Daily Emotion-20200214144103.csv"
mw_mood <- read_csv(mw_mood_fp)
```


# cleaning MW daily emo
```{r}
mw_mood_clean <-
  mw_mood %>%
  rename(
    User_ID = 
      `User Id`,
    ELS_ID =
      `First Name`,
    triggerdate =
      `Trigger Date`,
    triggertime =
      `Trigger Time`,
    DeviceOS = 
      `Device OS`,
    DailyEMO_sad = 
      `Since the last prompt, please indicate the extent to which you felt sad`,
    DailyEMO_annoyedorangry = 
      `Since the last prompt, please indicate the extent to which you felt annoyed/angry`,
    DailyEMO_grouchyorcranky =
      `Since the last prompt, please indicate the extent to which you felt grouchy/cranky`,
    DailyEMO_worried =
      `Since the last prompt, please indicate the extent to which you felt worried`,
    DailyEMO_anxious =
      `Since the last prompt, please indicate the extent to which you felt anxious`,
    DailyEMO_happy =
      `Since the last prompt, please indicate the extent to which you felt happy`,
    DailyEMO_cheerful =
      `Since the last prompt, please indicate the extent to which you felt cheerful`,
    DailyEMO_excited =
    `Since the last prompt, please indicate the extent to which you felt excited`,
    DailyEMO_energetic =
      `Since the last prompt, please indicate the extent to which you felt energetic`,
    DailyEMO_bored =
      `Since the last prompt, please indicate the extent to which you felt bored`,
    DailyEMO_want_other_people =
      `Since the last prompt, please indicate the extent to which you want to be with other people`,
    DailyEMO_want_alone =
      `Since the last prompt, please indicate the extent to which you want to be alone`,
    DailyEMO_who_are_you_with =
      `Right now, who are you with?`,
    DailyEMO_facetoface_YesorNo =
      `Right now, are you having a face-to-face conversation?`,
    DailyEMO_digital_convo_YesorNo =
      `Right now, are you having a real-time digital (phone, text, Facebook, video) conversation?`
  ) %>%
  mutate(
    User_ID = as.factor(User_ID)
  ) %>%
  select(-`Response Id`)
```

# pulling out only "c"
```{r}
mw_mood_c <-
  mw_mood_clean %>%
  mutate(
    porc =
      ifelse(str_detect(ELS_ID, "c"),
             "C",
             "P"
             )
    )

mw_mood_c <-
  mw_mood_c %>%
  filter(
    porc == "C"
  )

mw_mood_c_clean <-
  mw_mood_c %>%
  mutate(
    ELS_ID =
      str_remove_all(ELS_ID,
                     "[c]")
    ) %>%
  mutate(ELS_ID =
           as.numeric(ELS_ID)
         ) %>%
  select(
    ELS_ID,
    triggerdate,
    triggertime,
    starts_with("Daily")
  ) %>%
  mutate(
    ELS_ID =
      as.factor(ELS_ID)
  )
```

# if I need to control for mean moods
```{r}
mw_mood_c_collapse <-
  mw_mood_c_clean %>% 
  group_by(ELS_ID) %>%
   summarise(DailyEMO_sad_mean = 
               mean(DailyEMO_sad, na.rm=TRUE),  
             DailyEMO_annoyorang_mean = 
               mean(DailyEMO_annoyedorangry, na.rm=TRUE), 
             DailyEMO_grouchorcrank_mean = 
               mean(DailyEMO_grouchyorcranky, na.rm=TRUE),
             DailyEMO_worried_mean =
               mean(DailyEMO_worried, na.rm = TRUE),
             DailyEMO_anx_mean =
               mean(DailyEMO_anxious, na.rm = TRUE),
             DailyEMO_happy_mean = 
               mean(DailyEMO_happy, na.rm=TRUE), 
             DailyEMO_cheerful_mean = 
               mean(DailyEMO_cheerful, na.rm=TRUE), 
             DailyEMO_excited_mean = 
               mean(DailyEMO_excited, na.rm=TRUE), 
             DailyEMO_energetic_mean =
               mean(DailyEMO_energetic, na.rm=TRUE),
             DailyEMO_bored_mean =
               mean(DailyEMO_bored, na.rm=TRUE),
             DailyEMO_want_other_people_mean = 
               mean(DailyEMO_want_other_people, na.rm=TRUE), 
             DailyEMO_want_alone_mean = 
               mean(DailyEMO_want_alone, na.rm=TRUE)
             )
```

### parsing dates
```{r parsing dates}
mw_mood_c_clean <-
  mw_mood_c_clean %>%
  mutate(
    date1 = parse_date(triggerdate, "%d/%m/%y"),
    date2 = parse_date(triggerdate, "%d/%m/%Y"),
    date1 = ymd(date1),
    date2 = ymd(date2),
    date1 = as.character(date1),
    date2 = as.character(date2)
  )
```

## creating day and week variables
```{r}
# creating day and week variables
mw_mood_c_clean <-
  mw_mood_c_clean %>%
  mutate(
    date = date1,
    date = ifelse(
      !is.na(date2),
      date2,
      date
      )
    ) %>%
  mutate(
    day1 = day(date1),
    day2 = day(date2),
    wday1 = wday(date1, label = TRUE),
    wday2 = wday(date2, label = TRUE)
  ) %>%
  mutate(
    day1 = as.character(day1),
    day2 = as.character(day2),
    wday1 = as.character(wday1),
    wday2 = as.character(wday2)
  ) %>%
  mutate(
    day = day1,
    day = ifelse(
      !is.na(day2),
      day2,
      day
    ),
    week = wday1,
    week = ifelse(
      !is.na(wday2),
      wday2,
      week
      )
    ) %>%
  dplyr::select(
    -c(day1:wday2)
  )

# day num
mw_mood_c_clean_dayorder  <-
mw_mood_c_clean %>%
  left_join(mw_mood_c_clean %>% 
                        group_by(ELS_ID) %>% 
                        mutate(dayorder=order(date))) %>%
  mutate(
    week = as.factor(week)
  ) %>%
  ungroup()
```
# dealing with not enough or missing data  

if there are fewer than 2 days of data, I'm removing the subject
```{r not enough data}
# how many days per individual?
mw_mood_c_clean_dayorder <-
  mw_mood_c_clean_dayorder %>%
  group_by(ELS_ID) %>%
  mutate(
    data_num_obs =
      max(dayorder)
  ) %>%
  ungroup()

# filter those who have more than 1 day
mw_mood_filter  <-
  mw_mood_c_clean_dayorder %>%
  group_by(ELS_ID) %>%
  filter(
    data_num_obs > 1
  ) %>%
  dplyr::select(
    -c(
    triggerdate, date1, date2, day
    )
    ) %>%
  ungroup()
```

# averaging within day
just to reduce amount levels of the data to be comparable to sleep
```{r}
mw_mood_filter <-
  mw_mood_filter %>% 
  group_by(ELS_ID, dayorder) %>%
   summarize(DailyEMO_sad = 
               mean(DailyEMO_sad, na.rm=TRUE),  
             DailyEMO_annoyorang = 
               mean(DailyEMO_annoyedorangry, na.rm=TRUE), 
             DailyEMO_grouchorcrank = 
               mean(DailyEMO_grouchyorcranky, na.rm=TRUE),
             DailyEMO_worried =
               mean(DailyEMO_worried, na.rm = TRUE),
             DailyEMO_anx =
               mean(DailyEMO_anxious, na.rm = TRUE),
             DailyEMO_happy = 
               mean(DailyEMO_happy, na.rm=TRUE), 
             DailyEMO_cheerful = 
               mean(DailyEMO_cheerful, na.rm=TRUE), 
             DailyEMO_excited = 
               mean(DailyEMO_excited, na.rm=TRUE), 
             DailyEMO_energetic =
               mean(DailyEMO_energetic, na.rm=TRUE),
             DailyEMO_bored =
               mean(DailyEMO_bored, na.rm=TRUE),
             DailyEMO_want_other_people = 
               mean(DailyEMO_want_other_people, na.rm=TRUE), 
             DailyEMO_want_alone = 
               mean(DailyEMO_want_alone, na.rm=TRUE)
             )
```

# data viz

`DailyEMO_sad`, `DailyEMO_annoyorang`,   
`DailyEMO_grouchorcrank`, `DailyEMO_worried`,  
`DailyEMO_anx`, `DailyEMO_bored`, `DailyEMO_want_alone`,
`DailyEMO_happy`, `DailyEMO_cheerful`, `DailyEMO_excited`,  
`DailyEMO_energetic`, `DailyEMO_want_other_people`
```{r sad}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_sad,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r annoyed}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_annoyorang,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r grouchy}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_grouchorcrank,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r worried}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_worried,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r anxious}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_anx,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r bored}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_bored,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r want alone}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_want_alone,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r happy}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_happy,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r cheerful}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_cheerful,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r excited}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_excited,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r energetic}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_energetic,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```

```{r wants other people}
mw_mood_filter %>%
  ggplot(aes(dayorder, DailyEMO_want_other_people,
        group = ELS_ID)) +
      geom_line(alpha = 1/3) +
  facet_wrap(~ELS_ID)
```


```{r}
write_csv(mw_mood_filter,"~/Box/mw_mood_daily.csv")
```

